---
title: "Advent of Code 2025"
format: gfm
editor: source
execute:
  echo: false
  warning: false
  message: false
fig-dpi: 300
---

```{r}
library(tidyverse)
library(extrafont)
library(ggstar)

aoc_green <- "#009900"
aoc_gold <- "#ffff66"
github_grey <- "#0d1117"

# AoC-ish plot theme that blends into the GitHub dark mode
aoc_theme <- 
  theme(
    panel.background = element_rect(fill = NA, colour = NA, linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_rect(fill = github_grey, colour = github_grey),
    plot.title = element_text(colour = "white", size = 20, family = "Consolas"),
    strip.background = element_rect(fill = NA, colour = NA),
    strip.text = element_text(size = 16, colour = aoc_green),
    axis.text = element_text(size = 12, colour = "white", family = "Consolas"),
    axis.title = element_text(size = 14, colour = "white", family = "Consolas"),
    axis.line = element_line(colour = "white", linewidth = .5),
    axis.ticks = element_line(colour = "white", linewidth = .5),
    legend.position = "none"
  )

past_performance <- 
  data.frame(
    year = factor(2020:2025),
    stars = c(31, 33, 30, 40, 31, 14),
    total_stars = c(rep(50, 5), 24)
  ) %>%
  mutate(percent_stars = stars / total_stars * 100)
```

This repo contains my solutions for the [Advent of Code 2025](https://adventofcode.com/2025) in `R`. 
Note that I don't share my inputs, as requested in the [AoC About section](https://adventofcode.com/2025/about).

Here's my past performance (and this year's performance, so far). Note that in 2025, there were 12 instead of 25 puzzles (i.e., 24 instead of 50 gold stars to earn). This is why I display the performance as % stars earned.

```{r}
#| label: performance
#| fig.height: 2.5
past_performance %>% 
  ggplot(aes(x = year, y = percent_stars)) +
  geom_star(fill = aoc_gold, size = 5) +
  geom_path(colour = aoc_gold, group = 1) +
  scale_y_continuous(
    limits = c(min(past_performance$percent_stars) - 5, max(past_performance$percent_stars) + 5)
  ) +
  labs(y = "% stars earned") +
  aoc_theme
```

## TOC

Here is a table of contents showing you what each day was about, along with the run time. 
It's a crude measure, where I simply calculate the difference in `Sys.time()` at the beginning and the end of part 1, and the beginning and end of part 2. 
That means in some occasions, the runtime for part 2 will be much shorter because part of the code needed to solve part 2 was already run in part 1. 
Reading in the data (and sometimes, basic preprocessing that has nothing to do with the puzzle solution) is excluded from the run time calculations.

| Day | Title                   | Run Time            |
|---------------|---------------|---------------|
| 1   | Secret Entrance        | 0.30 / 120.91 ms     |
| 2   | Gift Shop        | 2.97 s / 4.75 mins :-(     |
| 3   | Lobby        | 15.11 / 15.43 ms    |
| 4   | Printing Department        | 0.19 / 3.63 s    |
| 5   | Cafeteria        | 86.01 / 84.8 ms    |
| 6   | Trash Compactor        | 6.21 / 59.7 ms    |
| 7   | Laboratories        | 7.27 / 73.42 ms    |
